Missing The Point:

### 2020-07-15 ICML/MissingThePoint.Rmd
Add take-away slide to ARTEMISS presentation

---
title: "Missing The Point:"
subtitle: "Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne I. Oberman, Stef van Buuren, Gerko Vink"
institute: "Utrecht University"
date: "17-07-2020" #(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

With iterative imputation, the validity of the inferences relies on algorithmic convergence.

```{r non-conv, echo=FALSE, fig.width=10, fig.height = 6.2, message=FALSE, warning=FALSE}
# set-up env
library(dplyr)
library(ggplot2)
library(plotly)
load("example_diagnostics.Rdata")

@@

# alternatively, use the actual data
# run 1.Execute.R until imputing anything, than run RunSimulationOnce.R until after amputing the data
# with the 'amps' object
library(mice)
load("../traceplotdata.Rdata")
set.seed(12)
d75 <- amps[[4]] %>% mice(maxit=10, printFlag = FALSE)
#plot(d75)
p <- gg.mids(d75, x="Y")[[1]] 
p <- p + 
  scale_color_manual(values = paint5) + 
  scale_x_continuous(breaks = 1:10) + 
  labs(
    #title = "Traceplot", 
    y = "Chain mean", 
    colour = "Imp.")

p %>% ggplotly()
```

@@

When imputing an incomplete multivariate normal set ( $n_{\rm obs}=1000$, $m=5$, $n_{\rm sim} = 1000$), we obtain valid regression estimates after 5 to 10 iterations. <!-- Multivariate normal set ( $n = 1000$ ), 'amputed' with a varying proportion of incomplete cases, and 'imputed' with a varying number of iterations. Depicted are the bias and coverage rate of an estimated regression coefficient ( $n_{\rm sim} = 1000$ ). -->

@@

---
# Thanks!

Read more: [github.com/hanneoberman/MissingThePoint](https://github.com/hanneoberman/MissingThePoint).

Or follow my updates on Twitter: [@hioberman](https://twitter.com/hioberman)

.footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).]

>>>>

---
title: "Missing The Point:"
subtitle: "Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne I. Oberman, Stef van Buuren, Gerko Vink"
institute: "ICML ARTEMISS workshop 2020"
date: "17-07-2020" #(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

@@

With iterative imputation, the validity of inferences relies on algorithmic convergence. Signs of non-convergence (i.e., non-mixing, trending) are typically identified through visual inspection.

<!-- .footer[Missing The Point | Oberman, van Buuren, Vink | 2020 | github.com/hanneoberman/MissingThePoint] -->

```{r non-conv, echo=FALSE, fig.width=10, fig.height = 5.2, message=FALSE, warning=FALSE}
# set-up env
library(dplyr)
library(ggplot2)
library(plotly)
library(mice)
#load("example_diagnostics.Rdata")
load("../traceplotdata.Rdata")

@@

# alternatively, use the actual data
# run 1.Execute.R until imputing anything, than run RunSimulationOnce.R until after amputing the data
# with the 'amps' object
set.seed(12)
d75 <- amps[[4]] %>% mice(maxit=10, printFlag = FALSE)
#plot(d75)
p <- gg.mids(d75, x="Y")[[1]] 
p <- p + 
  scale_color_manual(values = paint5) + 
  scale_x_continuous(breaks = 1:10) + 
  labs(
    #title = "Traceplot", 
    x = "Iteration",
    y = "Chain mean", 
    colour = "Imp.")

ggplotly(p)

```
---

@@

When imputing an incomplete multivariate normal set ( $n_{\rm obs}=1000$, $n_{\rm sim} = 1000$), we obtain valid regression estimates after 5 to 10 iterations. <!-- Multivariate normal set ( $n = 1000$ ), 'amputed' with a varying proportion of incomplete cases, and 'imputed' with a varying number of iterations. Depicted are the bias and coverage rate of an estimated regression coefficient ( $n_{\rm sim} = 1000$ ). -->

@@

---
# Take-away

We conclude that&mdash;in the cases considered&mdash;it never hurts to iterate longer, but such calculations hardly bring added value.

Read more on [github.com/hanneoberman/MissingThePoint](https://github.com/hanneoberman/MissingThePoint). Or follow my updates through Twitter [@hioberman](https://twitter.com/hioberman).

.pull-left[
<br> <br> <br>
<br> <br> <br>
<br> <br> <br>
Upcoming...
]

.pull-right[
<img src="https://raw.githubusercontent.com/gerkovink/shinyMice/2dae400d9d23ede7bec031c1501acdc681eab13c/ThesisProposal/Figures/logo.png" width="420">
]

<!-- .footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).] -->
